---
title: "Calculating Coefficients of the one-step predictors with the Innovations Algorithm IA()"
output: pdf_document
author: "Daniela Stängle"
---
```{r setup, include=FALSE}
library(devtools)
load_all(".")
knitr::opts_chunk$set(tidy = TRUE,echo = TRUE)
```
## Introduction
The Innovations Algorithm is a recursive algorithm that can be applied to stationary time series \(\{X_t\}\) with
zero-mean and finite second moments. The algorithm gives the coefficients of the one-step predictors \(\hat{X}_1,\hat{X}_2\dots\), which can be calculated recursively once the coefficients have been determined.
Additionally the algorithm calculates the mean squared errors \(\nu_i\).  

## Example 1: Basic Usage 
To apply the algorithm to a time series X, we call the function `IA()` with parameter X
```{r}
X <- stats::rnorm(5, mean = 0, sd = 1)
out <- zeitreihen::IA(X)
```
We obtain as output the vector `nu`, which contains the mean squared errors, 
```{r}
out$nu
```

the matrix \(\Theta_n=\begin{pmatrix}0&0&0&0&0\\ \theta_{11}&0&0&0&0\\ \theta_{22}&\theta_{21}&0&0&0\\
\theta_{33}&\theta_{32}&\theta_{31}&0&0 \\ \theta_{44}&\theta_{43}&\theta_{42}&\theta_{41}&0\\\end{pmatrix}\),
\(\\\)

```{r}
out$theta
```
and the vector of coefficients, `coeffs` =  \(\begin{pmatrix}\theta_{41}&\theta_{42}&\theta_{43}&\theta_{44}\end{pmatrix}\),
```{r}
out$coeffs
```
which can be used to calculate the one-step predictor \(\hat{X}_{n+1}\) via \(\newline\)
\(\hat{X}_{n+1}=\begin{cases} 0,& \text{if } n= 0,\\ \sum\limits_{j=1}^n\theta_{nj}\left(X_{n+1-j}-\hat{X}_{n+1-j}\right),& \text{if } n= 1,2,\dots,\end{cases}.\newline\)
For example the coefficient \(\theta_{41}\) is the coefficient before the difference of the value of the time series \(\{X_t\}\) at time \(t=4\) and its predictor \(\hat{X_4}\). 

## Example 2: Using the optional parameter `max_lag`
Additionally, the parameter `max_lag` (per default the length of the time series) can be set, to indicate till where the \(\theta\) 's should be calculated.
```{r}
X <- stats::rnorm(100, mean = 0, sd = 1)
out <- zeitreihen::IA(X, max_lag = 6)
out$nu
out$theta
```

## Example 3: Using incorrect values
There are several inputs which are not allowed and give the following warnings:\(\newline\)
1. If the time series contains values which are `NaN` or `Inf`:
```{r,error=TRUE}
X <- NaN + stats::rnorm(10, mean = 0, sd = 1)
Y <- c(0,0,0,Inf)
out <- zeitreihen::IA(Y)
out <- zeitreihen::IA(X)
```
\(\newline\)
2. If the time series contains values which are not numeric or complex values:
```{r, error=TRUE}
X <- "2.01"
out <- zeitreihen::IA(X)
```
\(\newline\)
3. If the time series is not an atomic vector:
```{r,error= TRUE}
X <- list(m=matrix(1,nrow = 10,ncol = 5), "test")
out <- IA(X)
```
\(\newline\)
4. If the `max_lag` is smaller than 3, not an integer or bigger than the length of the time series:
```{r,error=TRUE}
X <- stats::rnorm(10,0,1)
out <- zeitreihen::IA(X,2)
zeitreihen::IA(X,3.14)
zeitreihen::IA(X,51)
```


