---
title: "IA"
author: "Daniela Stängle"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{IA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(devtools)
load_all(".")
knitr::opts_chunk$set(tidy = TRUE, echo = TRUE)
```

## Calculating Coefficients of the one-step predictors with the Innovations Algorithm IA()
### Introduction
The Innovations Algorithm is a recursive algorithm that can be applied to stationary time series \(\{X_t\}\) with
zero-mean and finite second moments. The algorithm gives the coefficients of the one-step predictors \(\hat{X}_1,\hat{X}_2\dots\), which can be calculated recursively once the coefficients have been determined.
Additionally the algorithm calculates the mean squared errors \(\nu_i\).  

### Example 1: Basic Usage 
To apply the algorithm to a time series X, we call the function `IA()` with parameter X. Every time we call upon the algorithm, a warning will be given that the user has to check if the time series is stationary.
```{r }
X <- stats::rnorm(5, mean = 0, sd = 1)
out <- IA(X)
```
We obtain as output the vector `nu`, which contains the mean squared errors, 
```{r}
out$nu
```

the matrix \begin{equation*}\Theta_5=\begin{pmatrix}0&0&0&0&0\\ \theta_{11}&0&0&0&0\\ \theta_{22}&\theta_{21}&0&0&0\\
\theta_{33}&\theta_{32}&\theta_{31}&0&0 \\ \theta_{44}&\theta_{43}&\theta_{42}&\theta_{41}&0\\\end{pmatrix},\end{equation*}

```{r}
out$theta
```
and the vector of coefficients, `coeffs` =  \(\begin{pmatrix}\theta_{41}&\theta_{42}&\theta_{43}&\theta_{44}\end{pmatrix}\),
```{r}
out$coeffs
```
which can be used to calculate the one-step predictor \(\hat{X}_{5}\) via \(\newline\)
\begin{equation*}\hat{X}_{n}=\begin{cases} 0,& \text{if } n= 1,\\ \sum\limits_{j=1}^{n-1}\theta_{n-1,j}\left(X_{n-j}-\hat{X}_{n-j}\right),& \text{if } n= 2,3,4,5\end{cases} \end{equation*}
For example the coefficient \(\theta_{41}\) is the coefficient before the difference of the value of the time series \(\{X_t\}\) at time \(t=4\) and its predictor \(\hat{X_4}\). 

What happens:

* It is checked if `X` and `max_lag` have the required format.
* Allocation of a numeric vector `nu` of length `max_lag` to store the mean square errors $v_0,\dots,v_{max\_lag-1}$.
* Computation of the autocovariances ($\kappa(i,j)$ is equal to the autocovariance of X with $h=i-j$) of the time series with `sample_ACVF(X,h)` for `h` from 0 to the biggest difference `max_lag`-1.
* Initializing by assigning $v_0$ the autocovariance for `h` equal zero and allocating a numeric matrix with `max_lag` rows and columns.
* Calculating the coefficient $\theta_{11}$ and the mean squared error $v_1$ separately. Afterwards recursively calculating the coefficients $\theta_{n,n-k}$ and $v_k$ from $k=0$ to $n-1$ using the equation \begin{equation*}\theta_{n, n-k} = \nu_k^{-1} \left( \kappa(n+1, k+1) - \displaystyle{\sum_{j=0}^{k-1} \theta_{k, k-j} \theta_{n, n-j} \nu_j} \right), \quad 0 \leq k < n\end{equation*}
and \begin{equation*}\nu_n = \kappa(n+1, n+1) - \displaystyle{\sum_{j=0}^{n-1} \theta_{n, n-j}^2 \nu_j}.\end{equation*}


Required Format:

* `X` is an atomic vector of length bigger one, which has numeric or complex values and can't contain NA, Inf or NaN values.
* `max_lag` must be an integer bigger one, can't be NA, Inf or NaN and can't be bigger than the length of `X`.


### Example 2: Using the optional parameter `max_lag`
Additionally, the parameter `max_lag` (per default the length of the time series) can be set, to indicate till where the \(\theta\) 's should be calculated.
```{r warning=FALSE}
X <- stats::rnorm(100, mean = 0, sd = 1)
out <- zeitreihen::IA(X, max_lag = 6)
out$nu
out$theta
```


### Example 3: Using incorrect values
There are several inputs which are not allowed. Here are a few examples of the error messages:\(\newline\)
1. If the time series contains values which are `NaN` or `Inf`:
```{r,error=TRUE}
X <- NaN + stats::rnorm(10, mean = 0, sd = 1)
Y <- c(0, 0, 0, Inf)
out <- zeitreihen::IA(Y)
out <- zeitreihen::IA(X)
```
\(\newline\)
2. If the time series contains values which are not numeric or complex values:
```{r, error=TRUE}
X <- c("2.01", 1, 0)
out <- zeitreihen::IA(X)
```
\(\newline\)
3. If the time series is not an atomic vector:
```{r,error= TRUE}
X <- list(m = matrix(1, nrow = 10, ncol = 5), "test")
out <- IA(X)
```
\(\newline\)
4. If the `max_lag` is smaller than 2, not an integer or bigger than the length of the time series:
```{r,error=TRUE}
X <- stats::rnorm(10, 0, 1)
out <- zeitreihen::IA(X, 1)
zeitreihen::IA(X, 3.14)
zeitreihen::IA(X, 51)
```


