---
title: "Generate a synthetic Time Series with make_AR()"
author: "Alexander Klaus"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{make_MA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The `make_AR()` function generates a synthetic time series following an AR(p) (AutoRegressive of order p) model. This function allows you to create time series data based on specified AR coefficients and white noise variance.

## Example 1: Basic Usage of the Function

First, specify how long the time series should be:

```{r}
#library(zeitreihen)

t <- 100
```

Next, determine the coefficients of the AR(p) process, where `p` coefficients must be passed:

```{r}
phi <- 0.5
```

Now, generate the synthetic time series that follows the AR(p) model with the coefficients defined above:

```{r}
zeitreihen::make_AR(t, phi)
```

The output is a numeric or complex atomic vector of length `t` containing data generated by the AR model with hand over parameter. The output values corresponds to the sequential observations from 1 to `t`.

What is happening inside the function:

* It is checked within the function whether the inputs are in the correct format.
* An atomic vector $X$ of length `t` is created. If a vector `start` is passed, `X` contains this vector as values of the first `length(start)` entries.
* `t` normally distributed values with mean 0 and standard deviation `sigma` are generated.
* Iterative calculation of $X$ from `p + 1` to `t` with loop variable i:
    * Consider the vector, which stores the values from $X_{i - 1}$ ... $X_{i - q}$.
    * The dot product of the vector mentioned above and the vector phi is then calculated.
    * $Z[i]$ is added to the dot product.
    * And the previous calculation is stored in X[i].
* Finally, the values are returned.

## Example 2:
