---
title: "plot"
author: "Philip Kastura-Sahl"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, plot() setup, include = FALSE}
library(zeitreihen)

set.seed(742)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  fig.height = 6,
  fig.width = 6,
  tidy = TRUE
)
```

## the plotting function `plot()`
The periodogram of \(\{x_1, \dots, x_n\}\) is the function \(I_n(\lambda) = \frac{1}{n} \left| \sum_{t=1}^{n} x_t e^{-it \lambda} \right|^2\)
\
The `plot(X, n)`-function provides a spectral density plot of a time series. The function takes a time series as input and returns a plot of the spectral density.
```{r, plot() basic usage}
set.seed(742)
num_samples <- 10
X <- stats::rnorm(30)
zeitreihen::plot(X, num_samples)
```

### Adjusting the samples
By default, the plot takes `n`-samples from the interval of \((0, \pi]\) and evaluates the spectral density across it using the `periodogram()` function. To get a more detailed.
```{r, plot() adjust sample size}
num_samples <- 250
X <- stats::rnorm(30)
zeitreihen::plot(X, num_samples)
```

### Adjusting the interval
The `plot()` function also allows you to adjust the interval over which the spectral density is calculated by modifying the `from` and `to` parameters. By default, the interval is \((0, \pi]\), but you can specify a different interval that best suits your analysis. \
\
For example, to evaluate the spectral density over the interval \((\pi/2, \pi]\):
```{r, plot() adjusting interval}
num_samples <- 250
X <- stats::rnorm(30)
zeitreihen::plot(X, num_samples, from = pi / 2, to = pi)
```
By adjusting the `from` and `to` parameters, you can hone in on specific regions of interest in the frequency spectrum, allowing for more detailed spectral density analysis.

### Requirements
`X` should be a native vector consisting of either numeric (real numbers) or complex numbers. `X` should not include any values that are not finite such as `Inf`, `NaN` or `NA`.
```{r, plot() Example error with X, error = TRUE}
num_samples <- 10
X <- c(0, 1, 3, 4, Inf, NA)
zeitreihen::plot(X, num_samples)
```

`n` should be a whole number without any decimal component and should be a single value, not a vector.
```{r, plot() Example error with n, error = TRUE}
num_samples <- 10.42
X <- stats::rnorm(30)
zeitreihen::plot(X, num_samples)
```

### References
If you need more information, please refer to the package documentation or the following resources:

- [Introduction to Time Series and Forecasting](https://link.springer.com/book/10.1007/978-3-319-29854-2)
