---
title: "Using the Plotting function to Plot the estimated spectral density"
output: rmarkdown::html_vignette
author: "Philip Kastura-Sahl"
vignette: >
  %\VignetteIndexEntry{zeitreihen}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  tidy = TRUE,
  fig.width = 6,
  fig.height = 6,
  comment = "#>"
)
library(zeitreihen)
```

## `plot()`-Function for Spectral Density Plotting
The [**periodogram**](#periodogram) of \(\{x_1, \dots, x_n\}\) is the function
\[
  I_n(\lambda) = \frac{1}{n} \left| \sum_{t=1}^{n} x_t e^{-it\lambda} \right|^2
\]
\
The `plot(X, n)`-function provides a spectral density plot of a time series. The function takes a time series as input and returns a plot of the spectral density.
```{r, basic usage}
set.seed(742)
num_samples <- 10
X <- stats::rnorm(30)
zeitreihen::plot(X, num_samples)
```

## Adjusting the samples
The plot takes `n`-samples from the interval of \((0, \pi]\) and evaluates the spectral density across it using the `periodogram()` function. To get a more detailed
```{r, adjust sample size}
set.seed(742)
num_samples <- 250
X <- stats::rnorm(30)
zeitreihen::plot(X, num_samples)
```

## Requirements
- `X` should be a native vector consisting of either numeric (real numbers) or complex numbers. `X` should not include any values that are not finite such as `Inf`,`NaN` or `NA`.
```{r, Example error with X, error = TRUE}
set.seed(742)
num_samples <- 10
X <- c(0, 1, 3, 4, Inf, NA)
zeitreihen::plot(X, num_samples)
```
- `n` should be a whole number without any decimal component and should be a single value, not a vector.
```{r, Example error with n, error = TRUE}
set.seed(742)
num_samples <- 10.42
X <- stats::rnorm(30)
zeitreihen::plot(X, num_samples)
```

## References
If you need more information, please refer to the package documentation or the following resources:

- Introduction to Time Series and Forecasting: [Chapter 4.2 The Periodogram](https://link.springer.com/book/10.1007/978-3-319-29854-2)
