---
title: "zeitreihen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zeitreihen}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



# Zeitreihen: a time series Package
## Installation
To begin using `zeitreihen`, you first need to install the package. You can install it from CRAN:
```{r}
# install.packages("zeitreihen", repos = "https://cloud.r-project.org")
```

Or directly from GitHub (for the latest development version):
```{r}
# if(!require(devtools)) install.packages("devtools", repos = 'https://cloud.r-project.org')
# library(devtools)
# devtools::install_github("sirpiksel/rLang-s24")
```

```{r}
library(zeitreihen)
```

For this vignette, we'll use the dataset `tsdl`, by Rob Hyndman, Professor of Statistics at Monash University, Australia

```{r}
# devtools::install_github("FinYang/tsdl")
# library(tsdl)
# tsdl
```

## Conclusion

The `zeitreihen` package provides a powerful set of tools for data manipulation, visualization, and more. Whether you're working with magical creatures or mundane data, `zeitreihen` can help you get the job done efficiently.

For more detailed information on each function, refer to the function-specific documentation. Happy coding!

## References

If you need more information, please refer to the package documentation or the following resources:

- [R for Data Science](https://r4ds.had.co.nz/)
- [Advanced R](https://adv-r.hadley.nz/)
